Working title: Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-09-05 write_up/mixfishsim_ecolmod.tex
Edited to M&M

\usepackage{enumitem}

% For track-changes
%\usepackage[draft]{changes}
\usepackage[final]{changes} %% to turn off changes
\definechangesauthor[color = red]{JM}
\definechangesauthor[color = blue]{PD}
\definechangesauthor[color = green!40!black]{JJ}
\definechangesauthor[color = purple]{CM}

@@

\replaced[id = JJ]{Most fisheries}{Fishing} exploit\deleted[id=JJ]{s} a variety 
of spatially and temporally heterogeneous fish populations using
species-unselective gear. This can result in unintended, unwanted catch of low
quota or protected species and reducing these catches is crucial for
biological and economic sustainability of `mixed fisheries'. \\

\replaced[id = PD]{If fisheries are to avoid unwanted catch,}{To implement
	effective spatial measures to reduce discards} a good understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions and movement, and how fishers interact with different fish
populations. Because data on fish location at high temporal and spatial
resolutions is expensive and difficult to collect proxies are usually inferred
from scientific surveys or commercial catches with limited spatial and temporal
resolution. \\

To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal simulation model
incorporating: i) delay-difference population dynamics, ii) population movement
using Gaussian Random Fields to simulate patchy, heterogeneously distributed
populations, and iii) fishery dynamics for multiple fleet characteristics based
on species targeting under an explore-exploit strategy. This is implemented via
a mix of correlated random walk movement (for exploration) and learned
behaviour (for exploitation) phases of the fisheries.  \\

We simulated 50 years of fishing and used the results to draw inference on the
underlying community structures. We compared inference from commercial catch to
a simulated fixed-site sampling design and the true underlying community. We i)
establish the potential limitations of fishery-dependent data in providing a
robust picture of spatiotemporal distributions; and then ii) simulated an area
closure based on areas defined from the different data sources at a range of
temporal and spatial resolutions. We assessed the effectiveness of the closures
in reducing catches of a fish population. \\

Our framework allows users to explore the assumptions in modelling
observational data and evaluate the underlying dynamics of such approaches at a
fine spatial and temporal scale. We conclude from our simulations that
commercial data, while containing bias, provide a useful tool for managing
catches in mixed fisheries if applied at the correct spatiotemporal scale. \\

\end{abstract}

@@

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} appropriate
spatial scales for fisheries closures has been a challenge in the past but is
crucial to their success \citep{Costello2010, Dunn2016}. Inference is hampered
where spatial information is coarse due to low resolution reporting of
fisheries catch which is aggregated across larger gridded areas
\citep{Branch2005}. Further, if data does not allow identification of spatial
features it may lead to poorly sited closures which are ineffectual or have
unintended consequences. For example, increased benthic impact on previously
unexploited areas from the cod closure in the North Sea were observed without
the intended effect of reducing cod exploitation
\citep{Rijnsdorp2001,Dinmore2003}). \replaced[id=PD]{M}{Since then m}ore
refined spatiotemporal information has \added[id=PD]{since} become available
through the combination of logbook and Vessel Monitoring System (VMS) data
\citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more real-time
spatial management has been possible \citep[e.g.][]{Holmes2011}. However,
fishers establish favoured fishing grounds through an explore-exploit strategy
\citep{Bailey2018} where they search for areas with high catches and then use
experience to return to areas where they've experienced high catch in the past.
This leads to an inherently biased sampling where target species are
over-represented in catch as fishers exploit areas of high abundance.
\deleted[id=PD]{ Further, fishers generally only recorded landings (not catch)
	on a daily basis\todo{\added[id=JJ]{Its not clear what the problem is:
			landings collected on daily basis or landings recorded
			rather than catch}}.  This leads to questions about the
	validity of inference that can be drawn from landings data assigned to
	VMS activity pings.}\todo{\added[id=JJ]{This comes as a surprise: I
		thought this was going to be about
		discards}\added[id=PD]{Agree, have removed this to avoid
		confusion}}\\

\replaced[id=PD]{In order to understand the effect of spatiotemporal
	aggregation of data and fishery targeting we ask two fundamental
	questions regarding inference derived from observational data:}{ and
	use the results}\deleted[id=PD]{ from the fishing model:}

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time by use of a Gaussian spatial
process to define suitable habitat for different populations\replaced[id=PD]{.
	Precise locations of the fish being}{ and are} known
\added[id=PD]{directly} rather than inferred from sampling or commercial
catch\added[id=PD]{, we can use the population model to validate how inference
	from fisheries-dependent and fisheries independent sampling relates to
	the real community structure in a way we could not with real data}. We
ii) compare, at different spatial and temporal aggregations, the 'real
population' distributions to samples from fisheries-dependent and fisheries
independent catches to test if these are a true reflection of the relative
density of the populations. We then iii) simulate a fishery closure to protect
a species based on different spatial and
temporal data aggregations. \\

@@

%% the tikz figure
\clearpage
\input{Process_model.tex}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/pop_dist}
	\caption{Simulated spatial dynamics - the four populations abundance
		(log+1) at four time steps.
		}
	\label{fig:9}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width =0.5\linewidth]{./Plots/Closure_fishing_locations_yearly}
	\caption{Closure fishing locations based on annual closures with a
		coarse spatial resolution. Closure location can be seen in red
	in relation to a) before the closure fishing locations, b) after the
	closure fishing locations, c) species 3 habitat distribution}
	\label{fig:17}
\end{figure}

>>>>

\usepackage{enumitem}
\usepackage{comment}
% For track-changes
%\usepackage[draft]{changes}
\usepackage[final]{changes} %% to turn off changes
\definechangesauthor[color = red]{JM}
\definechangesauthor[color = blue]{PD}
\definechangesauthor[color = green!40!black]{JJ}
\definechangesauthor[color = purple]{CM}

@@

\replaced[id = JJ]{Most fisheries}{Fishing} exploit\deleted[id=JJ]{s} a variety 
of spatially and temporally heterogeneous fish populations using
species-unselective gear. This can result in unintended, unwanted catch of low
quota or protected species and reducing these catches is crucial for
biological and economic sustainability of `mixed fisheries'. \\
%%
\replaced[id = PD]{If fisheries are to avoid unwanted catch,}{To implement
	effective spatial measures to reduce discards} \replaced[id = CM]{an in-depth}{a good} understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions and movement, and how fishers interact with different fish
populations. Because data on fish location at high temporal and spatial
resolutions is expensive and difficult to collect proxies are usually inferred
from scientific surveys or commercial catches with limited spatial and temporal
resolution. \\ 
%%
To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal \added[id=CM]{event-based} simulation model
incorporating: i) delay-difference population dynamics, ii) population movement
using Gaussian Random Fields to simulate patchy, heterogeneously distributed
populations, and iii) fishery dynamics for multiple fleet characteristics based
on species targeting under an explore-exploit strategy. This is implemented via
a mix of correlated random walk movement (for exploration) and learned
behaviour (for exploitation) phases of the fisheries.  \\ 
%%
\replaced[id=CM]{Fifty}{We simulated 50} years of \added[id=CM]{sub-daily} fishing \added[id=CM]{was simulated} and used \deleted[id = CM]{the results}to draw inference on the
underlying community structures. We compared inference\added[id=CM]{s} \added[id=CM]{based on:}\deleted[id=CM]{from} commercial catch\added[id=CM]{,} \deleted[id=CM]{to}
a simulated fixed-site sampling \added[id=CM]{survey} design and the true underlying community. We i)
establish the potential limitations of fishery-dependent data in providing a
robust picture of spatiotemporal distributions; and then ii) simulated an area
closure based on areas defined from the different data sources at a range of
temporal and spatial resolutions. \deleted{We assessed the effectiveness of the closures
in reducing catches of a fish population. \\
%%
Our framework allows users to explore the assumptions in modelling
observational data and evaluate the underlying dynamics of such approaches at a
fine spatial and temporal scale.} We conclude from our \replaced[id=CM]{example framework application}{simulations} that
commercial data, while containing bias, provide a useful tool for managing
catches in mixed fisheries if applied at the correct spatiotemporal scale. [NEED A LITTLE MORE DETAIL OF THE RESULTS HERE - 2 MORE SENTENCES ON KEY TAKE-HOME RESULTS]\\
%%
\end{abstract}

@@

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} appropriate
spatial scales for fisheries closures has been a challenge in the past but is
crucial to their success \citep{Costello2010, Dunn2016}. Inference is hampered
where spatial information is coarse due to low resolution reporting of
fisheries catch which is aggregated across larger gridded areas
\citep{Branch2005}. Further, if data does not allow identification of spatial
features it may lead to poorly sited closures which are ineffectual or have
unintended consequences. For example, increased benthic impact on previously
unexploited areas from the cod closure in the North Sea were observed without
the intended effect of reducing cod exploitation
\citep{Rijnsdorp2001,Dinmore2003}). \replaced[id=PD]{M}{Since then m}ore
refined spatiotemporal information has \added[id=PD]{since} become available
through the combination of logbook and Vessel Monitoring System (VMS) data
\citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more real-time
spatial management has been possible \citep[e.g.][]{Holmes2011}. However,
fishers establish favoured fishing grounds through an explore-exploit strategy
\citep{Bailey2018} where they search for areas with high catches and then use
experience to return to areas where they've experienced high catch in the past.
This leads to an inherently biased sampling where target species are
over-represented in \added[id=CM]{the} catch as fishers exploit areas of high abundance.
\deleted[id=PD]{ Further, fishers generally only recorded landings (not catch)
	on a daily basis\todo{\added[id=JJ]{Its not clear what the problem is:
			landings collected on daily basis or landings recorded
			rather than catch}}.  This leads to questions about the
	validity of inference that can be drawn from landings data assigned to
	VMS activity pings.}\todo{\added[id=JJ]{This comes as a surprise: I
		thought this was going to be about
		discards}\added[id=PD]{Agree, have removed this to avoid
		confusion}}\\

\replaced[id=PD]{\replaced[id=CM]{To}{In order to} understand the effect of spatiotemporal
	aggregation of data and fishery targeting we ask two fundamental
	questions regarding inference derived from observational data:}{ and
	use the results}\deleted[id=PD]{ from the fishing model:}

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time by use of a Gaussian spatial
process to define suitable habitat for different populations\replaced[id=PD]{.
	Precise locations of the fish being}{ and are} known
\added[id=PD]{directly} rather than inferred from sampling or commercial
catch\added[id=PD]{, we can use the population model to validate how inference
	from fisheries-dependent and fisheries independent sampling relates to
	the real community structure in a way we could not with real data}. We
ii) compare, at different spatial and temporal aggregations, the \replaced[id=CM]{`}{'}real
population' distributions to samples from fisheries-dependent and fisheries
independent catches to test if these are a true reflection of the relative
density of the populations. We then iii) simulate a fishery closure to protect
a species based on different spatial and
temporal data aggregations. \\

@@

%% the tikz figure
\clearpage
\input{Process_model.tex}
\begin{comment}
\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/pop_dist}
	\caption{Simulated spatial dynamics - the four populations abundance
		(log+1) at four time steps.
		}
	\label{fig:9}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width =0.5\linewidth]{./Plots/Closure_fishing_locations_yearly}
	\caption{Closure fishing locations based on annual closures with a
		coarse spatial resolution. Closure location can be seen in red
	in relation to a) before the closure fishing locations, b) after the
	closure fishing locations, c) species 3 habitat distribution}
	\label{fig:17}
\end{figure}	
\end{comment}
